---
title: "PSY 423 Class 14"
author: "S. Budoff"
date: "2023-10-18"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r load}
library(tidyverse)
library(googlesheets4)
library(patchwork)
library(qrcode)

```

## Warmup Quiz

```{r qr}
code_link <- "https://forms.gle/9XRwL7KCKzPWEAMt8"
plot(qr_code(code_link))
print(code_link)
```


## Understanding Correlation

Correlation is a statistical concept that helps us understand the relationship between two CONTINUOUS variables.

Correlations Tell Us:

1) Magnitude of the relationship

2) Direction of the relationship

Pearson correlation measures the strength and direction of a linear relationship between two continuous variables.

It ranges from -1 (perfect negative correlation) to 1 (perfect positive correlation), with 0 indicating no linear correlation.

We call correlation r

## Hypotheses
In correlation, we typically test for the null and alternative hypotheses.
$$H_0 : \text{No correlation between Variable 1 and Variable 2}$$
$$H_a : \text{Correlation between Variable 1 and Variable 2}$$

## What do correlations look like?

```{r correlationPerfect}

df <- data_frame(X = rnorm(100, mean = 50)) %>%
  mutate(Y = X + 10)

plot1<-ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Perfect Correlation")+
  theme_minimal()

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 1)) %>%
  mutate(Y = X + 10 + rand)

plot2 <- ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Strong Correlation")+
  theme_minimal()

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 3)) %>%
  mutate(Y = X + 10 + rand)

plot3<-ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Weak Correlation")+
  theme_minimal()

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 5)) %>%
  mutate(Y = 45 + rand)

plot4<-ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("No Correlation")+
  theme_minimal()

library(patchwork)
plot1+plot2+plot3+plot4

```


## What factors effect r?

Outliers

Restriction of Range 

Nonlinearity

## Pearson Correlation Assumptions

1) Data is Randomly Selected and Representative 

2) Underlying distributions are approximately normal (and data are continuous)

3) Linear Relationship

4) No outliers

## Applying Correlation in Psychometrics


The field of study concerned with the measurement of psychological attributes and constructs.

Two fundamental concepts in psychometrics are reliability and validity, which play a crucial role when examining relationships between variables, such as in correlations.


## Reliability

Refers to the consistency and stability of measurements over time or across different situations.

In the context of correlations, reliable ensure that the relationship between variables is accurately captured.

Common methods for assessing reliability include: 
- test-retest reliability 
- Split-Half test 
- Cronbach's $\alpha$

## Test-Retest Reliability

Assesses the consistency of measurements by comparing results from the same test administered at two different points in time.

In correlation studies, if the measurements are not reliable, the observed relationships between variables may be due to measurement error rather than true associations.

## Split-Half Testing

Split-Half testing is a method used to assess the internal consistency reliability of a measurement instrument.

In this approach, a test or scale is divided into two halves, and the scores on each half are correlated.

The correlation coefficient is typically used to assess the degree to which both halves yield similar results.

## Advantages of Split-Half Testing

Provides a quick and straightforward way to estimate internal consistency.

Requires less time and resources compared to other reliability assessment methods.

Allows for an initial assessment of the reliability of a measurement instrument.

## Limitations of Split-Half Testing

The reliability estimate may be influenced by the specific way in which the test is divided into halves.

The reliability estimate may be lower than the true internal consistency, especially for short tests.

## Cronbach's Alpha ($\alpha$)

Cronbach's Alpha, also known as coefficient $\alpha$, is a widely used method to assess internal consistency reliability.

It is a more sophisticated and accurate measure than split-half testing and is suitable for a wide range of measurement instruments, including multi-item scales.

## Calculating Cronbach's $\alpha$

Cronbach's $\alpha$ is calculated based on the average intercorrelation among all items in a scale.
It assesses how closely items are related to one another, providing a more comprehensive measure of internal consistency.

## Interpretation of Cronbach's $\alpha$

Cronbach's $\alpha$ values range from 0 to 1.

A higher Cronbach's $\alpha$ indicates greater internal consistency, with values above 0.7 generally considered acceptable, and values above 0.8 are often preferred in psychological research.

## Advantages of Cronbach's $\alpha$

Provides a more accurate and comprehensive estimate of internal consistency.

Suitable for measuring the reliability of complex measurement instruments, including multi-item scales

## JAMOVI Time

## Validity

Refers to the extent to which a test or measurement instrument accurately assesses the construct it is intended to measure.

In correlation studies, validity is essential to ensure that the variables being studied truly represent the concepts of interest.

## Content Validity

Content validity involves examining the relevance and representativeness of the items in a test in relation to the construct being measured.

For correlations, content validity ensures that the variables being correlated are pertinent to the research question.

For example, the material on a comprehensive final exam

## Construct Validity

Construct validity assesses whether the test or measure accurately reflects the theoretical construct it is intended to represent.

In correlation studies, construct validity is vital to confirm that the variables being correlated indeed reflect the underlying psychological constructs.

For example if you are studying agoraphobia, you may like your measure to differentiate features of anxiety from features of arachnophobia 

## Criterion-Related Validity

Criterion-related validity evaluates the extent to which a test predicts or correlates with a relevant criterion.

When studying correlations, criterion-related validity helps establish that the variables being correlated have meaningful relationships with external criteria.

Typically, this is a comparison with a gold-standard or otherwise exisitng measure

For example, if you develope a new intelligence test comparing participant's scores with the classic IQ test

## Try to come up with your own examples:

- Content

- Construct

- Criterion

## The Interplay

Reliability and validity are interrelated. 

A reliable measure is a prerequisite for a valid measure, and a valid measure must also be reliable.

In correlation research, ensuring both reliability and validity is essential to draw accurate and meaningful conclusions about the relationships between variables.

## Conclusion

Reliability and validity are essential concepts in psychometrics when conducting correlation studies.

Reliability ensures that measurements are consistent and stable over time, while validity ensures that measurements accurately reflect the constructs of interest.

Balancing both reliability and validity is crucial for robust and meaningful correlations in psychological research.





