---
title: "PSY 423 Class 12"
author: "S. Budoff"
date: "2023-10-10"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r load}
library(tidyverse)
library(googlesheets4)
library(patchwork)
library(qrcode)

gs4_auth(email = "sabudoff@gmail.com")


url = "https://docs.google.com/spreadsheets/d/1sVKUiXkLEOgTMiaxK__h8opR8WmN4hJoHtP-hGvcf40/edit#gid=0"
df <- read_sheet(url, sheet = "Exam1")
df <- df %>%
  select(Score, Scantron, Doodles, `Final Score`) %>%
  mutate(`Final Score` = round(`Final Score`))
```

## Warmup Quiz

```{r qr}
code_link <- "https://forms.gle/h3ERZoEojYk8Jt2M7"
plot(qr_code(code_link))
print(code_link)
```

## Exam Distribution - Final Scores

```{r exam1}
ggplot(df, aes(x = `Final Score`)) +
  geom_histogram(bins = 5) +
  theme_minimal()
```

## Exam Distribution - Doodles

```{r exam2}
ggplot(df, aes(x = Doodles)) +
  geom_histogram(bins = 5) +
  theme_minimal()
```


## Exam Distribution - Scantron

```{r exam3}
ggplot(df, aes(x = Scantron)) +
  geom_histogram(bins = 5) +
  theme_minimal() 
```

## Exam Distribution - Take Home

```{r exam4}
ggplot(df, aes(x = Score)) +
  geom_histogram(bins = 5) +
  theme_minimal()
```

## JAMOVI Review - Exam Question


## 2-Factor ANOVA

Instead of one independent variable, we have 2!

For example, the height of a person may be related to their gender AND if they play basketball

In your groups think of some other scenarios like this where multiple independent variable relate to a dependent variable

## Why not just do 2 separate 1-Factor ANOVAs?


## Why not just do 2 separate 1-Factor ANOVAs?

We can study the INTERACTION between factors!

Study systemic variability coming from an additional factor

## The 2-Factor ANOVA tells you about

How Factor A relates to the dependent variable

How Factor B relates to the dependent variable

How Factor A and Factor B interact in relation to the dependent variable


## Example Set up

2+ levels of IV1 (e.g. Drug A, B, C)

2+ levels of IV2 (e.g. Low, High Dosage)

1 continuous DV (e.g. anxiety)

What are the hypothesis we are testing?


```{r fakedata}
set.seed(123)
group1 <- rnorm(10, mean=10, sd=2) 
group2 <- rnorm(10, mean=12, sd=3)
group3 <- rnorm(10, mean=14, sd=4)
dv <- c(group1, group2, group3)
iv1 <- rep(c("A","B","C"), each=10)
iv2 <- rep(c("High","Low"), times=15) 
df <- data.frame(dv=dv, iv1=iv1, iv2=iv2)
```

## Hypotheses

$$H_0: \mu_{A} = \mu_{B} = \mu_{C}$$

$$H_a: \text{at least one mean differs}$$

$$H_0: \mu_\text{Low} = \mu_\text{High}$$
$$H_a: \mu_\text{Low} \neq \mu_\text{High}$$

$$H_0: \text{No Interaction}$$
$$H_a: \text{Significant Interaction}$$

## Analysis

- Two-way ANOVA 

- Follow up tests

  - Main effects: Tukey's HSD

  - Interaction: Simple effects analysis



## Interaction Plot


```{r intplot}
# Run two-way ANOVA
aov_out <- aov(dv ~ iv1*iv2, data=df)

# Get Tukey HSD for main effects
# TukeyHSD(aov_out)

# Check interaction 
interaction.plot(iv1, iv2, dv)
```
The interaction plot shows the mean of the dependent variable (y-axis) for each combination of the levels of the two independent variables (x-axis).

## Interaction Plot Interpretations:

Parallel lines indicate no interaction between the IVs. The effects of one IV are consistent across all levels of the other IV.

Non-parallel lines indicate an interaction. The effect of one IV depends on the level of the other IV.

If the lines cross, this is a significant interaction. The IVs influence each other.

If the lines are not parallel but do not cross, the interaction may still be significant, but the IVs do not reverse the effects of each other.
Things to look for:

## Questions to ask when looking at Interaction Plots:

Are the lines roughly parallel? Or do they cross/diverge?

Does the range of the DV means differ across levels of one IV at different levels of the other IV?

Is there an extreme difference between one particular combination versus the others?

The interaction plot gives a visual depiction of the way the IVs are interacting. It supplements the statistical test of the interaction term in the ANOVA.

