---
title: "PSY 423 Class 14"
author: "S. Budoff"
date: "2023-10-18"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r load}
library(tidyverse)
library(googlesheets4)
library(patchwork)
library(qrcode)

```

## Warmup Quiz

```{r qr}
code_link <- "https://forms.gle/9XRwL7KCKzPWEAMt8"
plot(qr_code(code_link))
print(code_link)
```


## Understanding Correlation

Correlation is a statistical concept that helps us understand the relationship between two CONTINUOUS variables.

Correlations Tell Us:

1) Magnitude of the relationship

2) Direction of the relationship

Pearson correlation measures the strength and direction of a linear relationship between two continuous variables.

It ranges from -1 (perfect negative correlation) to 1 (perfect positive correlation), with 0 indicating no linear correlation.

We call correlation r

## Hypotheses
In correlation, we typically test for the null and alternative hypotheses.
$$H_0 : \text{No correlation between Variable 1 and Variable 2}$$
$$H_a : \text{Correlation between Variable 1 and Variable 2}$$

## What do correlations look like?

```{r correlationPerfect}

df <- data_frame(X = rnorm(100, mean = 50)) %>%
  mutate(Y = X + 10)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Perfect Correlation")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)


## What do correlations look like?

```{r correlationGood}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 1)) %>%
  mutate(Y = X + 10 + rand)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Strong Correlation")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)

## What do correlations look like?

```{r correlationweak}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 3)) %>%
  mutate(Y = X + 10 + rand)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Weak Correlation")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)

## What do correlations look like?

```{r correlationNone}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 5)) %>%
  mutate(Y = 45 + rand)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("No Correlation")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)


## What do correlations look like?

```{r correlationNeg}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 1)) %>%
  mutate(Y = -X + 10 + rand)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Negative Correlation")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)

## What factors effect r?

Outliers

Restriction of Range 

Nonlinearity

## Outliers

```{r outliers}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean = 5, sd = 5)) %>%
  mutate(i = 1:n(),
         Y = case_when(i < 100 ~ 45 + rand,
                       TRUE ~ 60 + rand),
         X = case_when(i < 100 ~ X,
                       TRUE ~ i-30))


ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Effect of Outliers")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)

## Restriction of range

```{r restriction1}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 1)) %>%
  mutate(Y = X + 10 + rand)


ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Strong Correlation")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)

## Restriction of range

```{r restriction2}

set.seed(1)
df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 1)) %>%
  mutate(Y = X + 10 + rand) %>%
  filter(X > 49, X < 50)


ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Effect of Restriction of Range")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)



## Nonlinearity

```{r nonlin}

set.seed(1)
df <- data_frame(X = rnorm(100, mean = 0),
                 rand = rnorm(100, mean=5, sd = 1)) %>%
  mutate(Y = X^2 + rand) 


ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Effect of Nonlinearity")+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(df$Y, df$X, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)


## Pearson Correlation Assumptions

1) Data is Randomly Selected and Representative 

2) Underlying distributions are approximately normal (and data are continuous)

3) Linear Relationship

4) No outliers

## What is the Pearson Correlation

It is the ratio between the covariance of two variables

It is a normalized measure of covariance, this is why it is always between -1 and 1!

We use a (one-sample) t-test to assess if there is a significant correlation because the Correlation Coefficient (r) follows a t-distribution when our assumptions are true

What is the theoretic mean from our $H_0$?



## Housing Data Examples


```{r house1}
data <- read.csv("/home/sam/Regis/PSY423/AmesHousing.csv")

ggplot(data, aes(y=SalePrice, x = Lot.Area)) +
  geom_point(alpha = 0.3) +
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Lot.Area, method = "pearson")

```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)

## Housing Data Examples


```{r house2}

ggplot(data, aes(y=SalePrice, x = Year.Built)) +
  geom_point(alpha = 0.3) +
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Year.Built, method = "pearson")
```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)


## Housing Data Examples


```{r house4}

ggplot(data, aes(y=SalePrice, x = Garage.Area)) +
  geom_point(alpha = 0.3) +
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Garage.Area, method = "pearson")
```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)

## Housing Data Examples


```{r house3}

ggplot(data, aes(y=SalePrice, x = Pool.Area)) +
  geom_point(alpha = 0.3) +
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Pool.Area, method = "pearson")
```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)


## Housing Data Examples


```{r house5}
data2 <- data %>%
  filter(Pool.Area > 0)


ggplot(data2, aes(y=SalePrice, x = Pool.Area)) +
  geom_point(alpha = 0.3) +
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data2$SalePrice, data2$Pool.Area, method = "pearson")
```

t(`r cor_test_result$parameter`) = `r cor_test_result$statistic`, p = `r  cor_test_result$p.value`, r = `r  cor_test_result$estimate`)


## Analysis Interpreting the Result

Interpret the result of the correlation test.

Key values the correlation coefficient, p-value, and the strength of correlation.


## Visualizing Correlation

It's often helpful to visualize the correlation.

Show a heatmap to visualize correlations between multiple variables.



