---
title: "PSY 423 Class 14"
author: "S. Budoff"
date: "2023-10-18"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r load}
library(tidyverse)
library(googlesheets4)
library(patchwork)
library(qrcode)

```

## Warmup Quiz

```{r qr}
code_link <- "https://forms.gle/i3aGWkWkrSUPk6Yn8"
plot(qr_code(code_link))
print(code_link)
```


## 2-Factor ANOVA Steps

- Two-way ANOVA 

- Follow up tests

  - Main effects: Tukey's HSD

  - Interaction: Simple effects analysis



## Interaction Plot - No Signifigance


```{r intplot1}

set.seed(123)
group1 <- rnorm(100, mean=10, sd=2) 
group2 <- rnorm(100, mean=10, sd=2)
group3 <- rnorm(100, mean=10, sd=2)
dv <- c(group1, group2, group3)
iv1 <- rep(c("A","B","C"), each=100)
iv2 <- rep(c("High","Low"), times=150) 
df <- data.frame(dv=dv, iv1=iv1, iv2=iv2)
# Run two-way ANOVA
aov_out <- aov(dv ~ iv1*iv2, data=df)
print(summary(aov_out))

# Get Tukey HSD for main effects
# TukeyHSD(aov_out)

# Check interaction 
interaction.plot(iv1, iv2, dv)
```

## Interaction Plot - Signifigance but no interaction


```{r intplot2}

set.seed(123)
group1 <- rnorm(100, mean=10, sd=2) 
group2 <- rnorm(100, mean=12, sd=2)
group3 <- rnorm(100, mean=14, sd=2)
dv <- c(group1, group2, group3)
iv1 <- rep(c("A","B","C"), each=100)
iv2 <- rep(c("High","Low"), times=150) 
df <- data.frame(dv=dv, iv1=iv1, iv2=iv2)
# Run two-way ANOVA
aov_out <- aov(dv ~ iv1*iv2, data=df)
print(summary(aov_out))

# Get Tukey HSD for main effects
# TukeyHSD(aov_out)

# Check interaction 
interaction.plot(iv1, iv2, dv)
```

## Interaction Plot - Signifigance but no interaction


```{r intplot3}
set.seed(123)
group1 <- rnorm(10, mean=10, sd=2) 
group2 <- rnorm(10, mean=12, sd=3)
group3 <- rnorm(10, mean=14, sd=5)
dv <- c(group1, group2, group3)
iv1 <- rep(c("A","B","C"), each=10)
iv2 <- rep(c("High","Low"), times=15) 
df <- data.frame(dv=dv, iv1=iv1, iv2=iv2)
# Run two-way ANOVA


# Get Tukey HSD for main effects
# TukeyHSD(aov_out)

# Check interaction 
interaction.plot(iv1, iv2, dv)
```




The interaction plot shows the mean of the dependent variable (y-axis) for each combination of the levels of the two independent variables (x-axis).

## Interaction Plot Interpretations:

Parallel lines indicate no interaction between the IVs. The effects of one IV are consistent across all levels of the other IV.

Non-parallel lines indicate a possible interaction. The effect of one IV depends on the level of the other IV.

If the lines cross, this is strong visual evidence for a significant interaction. The IVs influence each other.

It is also possible for lines to cross if the assumptions (like homoscedasticity) are violated!

## Questions to ask when looking at Interaction Plots:

Are the lines roughly parallel? Or do they cross/diverge?

Does the range of the DV means differ across levels of one IV at different levels of the other IV?

Is there an extreme difference between one particular combination versus the others?

The interaction plot gives a visual depiction of the way the IVs are interacting. It supplements the statistical test of the interaction term in the ANOVA.


## Understanding Correlation

Correlation is a statistical concept that helps us understand the relationship between two CONTINUOUS variables.

## Scenarios

Think of scenarios where correlation might be relevant. For example, how might temperature correlate with ice cream sales?

## What Do Correlations Tell Us?

## What Do Correlations Tell Us?

Magnitude of the relationship

Direction of the relationship

## What Do Correlations NOT Tell Us?

## What Do Correlations NOT Tell Us?

Causation!

## What is Causation then?

A relationship between variables where variations in one is the best possible explanation for the other

Causation has 2 Preconditions:

1) A significant **correlation**
2) A precise temporal relationship and theoretic explanation for this relationship



## Pearson Correlation

Pearson correlation measures the strength and direction of a linear relationship between two continuous variables.

It ranges from -1 (perfect negative correlation) to 1 (perfect positive correlation), with 0 indicating no linear correlation.



## Hypotheses
In correlation, we typically test for the null and alternative hypotheses.
$$H_0 : \text{No correlation between Variable 1 and Variable 2}$$
$$H_a : \text{Correlation between Variable 1 and Variable 2}$$

## What do correlations look like?

```{r correlationPerfect}

df <- data_frame(X = rnorm(100, mean = 50)) %>%
  mutate(Y = X + 10)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Perfect Correlation")+
  theme_minimal()

```
## What do correlations look like?

```{r correlationGood}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 1)) %>%
  mutate(Y = X + 10 + rand)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Strong Correlation")+
  theme_minimal()

```

## What do correlations look like?

```{r correlationweak}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 3)) %>%
  mutate(Y = X + 10 + rand)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Weak Correlation")+
  theme_minimal()

```
## What do correlations look like?

```{r correlationNone}

df <- data_frame(X = rnorm(100, mean = 50),
                 rand = rnorm(100, mean=5, sd = 5)) %>%
  mutate(Y = 45 + rand)

ggplot(df, aes(x=X, y=Y)) +
  geom_point() +
  ggtitle("Probabaly No Correlation")+
  theme_minimal()

```

## Housing Data Examples


```{r house1}
data <- read.csv("/home/sam/Regis/PSY423/AmesHousing.csv")

ggplot(data, aes(y=SalePrice, x = Lot.Area)) +
  geom_point(alpha = 0.3) +
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Lot.Area, method = "pearson")
print(cor_test_result)
```

## Housing Data Examples


```{r house2}

ggplot(data, aes(y=SalePrice, x = Year.Built)) +
  geom_point(alpha = 0.3) +
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Year.Built, method = "pearson")
print(cor_test_result)
```

## Housing Data Examples


```{r house3}

ggplot(data, aes(y=SalePrice, x = Pool.Area)) +
  geom_point(alpha = 0.3) +
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Pool.Area, method = "pearson")
print(cor_test_result)
```


## Housing Data Examples


```{r house4}

ggplot(data, aes(y=SalePrice, x = Garage.Area)) +
  geom_point(alpha = 0.3) +
  theme_minimal()

# Perform correlation test
cor_test_result <- cor.test(data$SalePrice, data$Garage.Area, method = "pearson")
print(cor_test_result)
```

## Analysis Interpreting the Result

Interpret the result of the correlation test.

Key values the correlation coefficient, p-value, and the strength of correlation.


## Visualizing Correlation

It's often helpful to visualize the correlation.

Show a heatmap to visualize correlations between multiple variables.



