---
title: "PSY 423 Class 6"
author: "S. Budoff"
date: "2023-09-11"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r load}
library(tidyverse)
library(googlesheets4)
library(patchwork)
library(qrcode)

gs4_auth(email = "sabudoff@gmail.com")

#Sort the integer values
coffee_levels <- c("Love it!", "It's good", "No opinion", "It's bad", "Ewwww!")
shoe_levels <- c("M 0-4.5 / W 0-6",
                 "M 5-6 / W 6.5-7.5",
                 "M 6.5-7.5 / W 8-9",
                 "M 8-9 / W 9.5-10.5",
                 "M 9.5-10.5 / W 11-12",
                 "M 11-12 / W 12.5-13.5",
                 "M 12.5+ / W14+")

url = "https://docs.google.com/spreadsheets/d/18Vj9uuD51ZHRds2Jqve91Cii_x46lJJl5ErrEwb4tKs/edit#gid=976217741"
df <- read_sheet(url, sheet = "Form Responses 1")
df <- df %>%
  mutate(Shoe_Size = factor(`What size shoe do you wear?`, levels=shoe_levels),
         HairColor = factor(`What color is your hair right now?`),
         Height = `How tall are you (in inches)?`,
         Coffee = factor(`How much do you like coffee?`, levels=coffee_levels)) %>%
  select(Shoe_Size, HairColor, Height, Coffee)
```

## Warmup Quiz

```{r qr}
code_link <- "https://forms.gle/gnhCP36HpQuTmnwa7"
plot(qr_code(code_link))
print(code_link)
```


## Confidence Intervals

A set/interval of numbers that contain the parameter we care about a specified percent of time

Usually 95%

If the parameter we care about is the mean, in frequentist statistics it is the interval we predict to contain the sample mean for 95 out of 100 experiments we conduct


## How do you think sample size will relate to confidence intervals?

```{r std}
# Create a data frame with different sample sizes
sample_sizes <- seq(2, 200, by = 10)
data <- data.frame(sample_size = sample_sizes)

# Generate means and standard errors for each sample size
data$mean <- sapply(sample_sizes, function(n) {
  x <- rnorm(n)  # Generate random data
  mean(x)
})

data$std_error <- sapply(sample_sizes, function(n) {
  x <- rnorm(n)  # Generate random data
  sd(x) / sqrt(n)  # Calculate standard error
})

# Create a plot
ggplot(data, aes(x = sample_size, y = mean, ymin = mean - 1.96 * std_error, ymax = mean + 1.96 * std_error)) +
  geom_point() +
  geom_errorbar(width = 0.2) +
  labs(
    x = "Sample Size",
    y = "Mean",
    title = "Effect of Sample Size on Confidence Interval Width"
  ) +
  theme_minimal()
```

## N's effect on CIs come rom the standard error of the mean

Increasing the sample size decreases the width of confidence intervals, because it decreases the standard error.

Standard error (SE) of estimate is a statistic of the stability and reliability of data. It is the variability across MULTIPLE samples

Standard error is used to estimate the standard deviation, which is the variability across a population, or in the case of a sample standard deviation across a SINGLE sample.

standard error:
$$\sigma_{\bar{x}} = \frac{s}{\sqrt{n}}$$


