---
title: "PSY 423 Class 6"
author: "S. Budoff"
date: "2023-09-11"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

```{r load}
library(tidyverse)
library(googlesheets4)
library(patchwork)
library(qrcode)

gs4_auth(email = "sabudoff@gmail.com")

#Sort the integer values
coffee_levels <- c("Love it!", "It's good", "No opinion", "It's bad", "Ewwww!")
shoe_levels <- c("M 0-4.5 / W 0-6",
                 "M 5-6 / W 6.5-7.5",
                 "M 6.5-7.5 / W 8-9",
                 "M 8-9 / W 9.5-10.5",
                 "M 9.5-10.5 / W 11-12",
                 "M 11-12 / W 12.5-13.5",
                 "M 12.5+ / W14+")

url = "https://docs.google.com/spreadsheets/d/18Vj9uuD51ZHRds2Jqve91Cii_x46lJJl5ErrEwb4tKs/edit#gid=976217741"
df <- read_sheet(url, sheet = "Form Responses 1")
df <- df %>%
  mutate(Shoe_Size = factor(`What size shoe do you wear?`, levels=shoe_levels),
         HairColor = factor(`What color is your hair right now?`),
         Height = `How tall are you (in inches)?`,
         Coffee = factor(`How much do you like coffee?`, levels=coffee_levels)) %>%
  select(Shoe_Size, HairColor, Height, Coffee)
```

## Warmup Quiz

```{r qr}
code_link <- "https://forms.gle/eSGk1eAhABeh6Whc6"
plot(qr_code(code_link))
print(code_link)
```

## The z and t distributions 

When $sigma$ is known

$$z = \frac{x - \mu}{\sigma} \implies \frac{x - \bar{x}}{\sigma} $$ 

When $sigma$ is unknown

$$t = \frac{x - \mu}{s_{\bar{x}}} \implies \frac{x - \bar{x}}{s_{\bar{x}}} = \frac{x - \bar{x}}{\frac{s}{\sqrt{n}}} $$ 

## Visualizing the t-distributions

In your groups doodle a Normal distribution, a t-distribution with df = 29, and a t-distribution with df = 2

## Visualizing the t-distributions

```{r tdists}

# Generate a sequence of x values
x <- seq(-4, 4, by = 0.01)
dfs <- c(1, 3, 10, 30)
# Create a data frame with x values and PDF values for different degrees of freedom
df <- data.frame(x = rep(x, times = length(dfs)), 
                 df = rep(dfs, each = length(x)),
                 pdf = dt(x, df = rep(dfs, each = length(x))))

# Create the plot
ggplot(df, aes(x = x, y = pdf, color = as.factor(df))) +
  geom_line() +
  geom_line(data = data.frame(x = x, pdf = dnorm(x)), aes(x = x, y = pdf), linetype = "dashed", color = "black") + # Add the standard normal PDF
  labs(title = "T-Distribution PDF with Different Degrees of Freedom", y = "Probability", x="standard deviation") +
  scale_color_discrete(name = "Degrees of Freedom") +
  theme_minimal() 
```


## Typese of t-tests

1) Single Sample

2) Independent Samples

3) Paired Samples (AKA: Dependent Samples)

## Single Sample

Used when we have a predicted mean ($\mu_0$) we would like to assess a single sample distribution from

Think of an experiment where this is applicable

Come up with the null and alternative hypothesis

What will be the degrees of freedom?

Try to write the test equation


## Single Sample

Used when we have a predicted mean ($\mu_0$) we would like to assess a single sample distribution from

$$H_0 : \bar{x} = \mu_0$$

$$H_a : \bar{x} \neq \mu_0$$

$$df = n-1$$

$$ t = \frac{\bar{x} - \mu_0}{s_{\bar{x}}} = \frac{\bar{x} - \mu_0}{\frac{s}{\sqrt{n}}}$$

## JAMOVI example

The average height for men in America is 69.1 inches, and for women it is 63.5 inches. The average of these values is 66 inches. 

Given our class height data, what is the null hypothesis?



## Independent Samples

Used when we have two samples and we would like to predict if they come from the same distribution

Think of an experiment where this is applicable

Come up with the null and alternative hypothesis

What is funny about the sample size now? Can you think of a math trick we have used before to deal with it?

What will be the degrees of freedom?

Try to write the test equation


## Independent Samples

Used when we have a predicted mean ($\mu_0$) we would like to assess a single sample distribution from

$$H_0 : \bar{X}_1 = \bar{X}_2 \text{   OR   } \mu_1 = \mu_2$$

$$H_a : \bar{X}_1 \neq \bar{X}_2  \text{   OR   } \mu_1 \neq \mu_2$$

$$df = n_1 + n_2 - 2$$

$$ t = \frac{\bar{X}_1 - \bar{X}_2}{SEM_{pooled}} = \frac{\bar{X}_1 - \bar{X}_2}{s_{pooled}\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}} $$

## JAMOVI example

They say that coffee makes you shorter, is this true in our class?

What is the null hypothesis for a two tailed test? How about for this myth as stated?

## Paired Samples

Used when we have two samples that are matched in some way and we would like to predict if their distribution has changed

Think of an experiment where this is applicable

Can you think of a math trick we have used before to see if there is a difference between pairs?

Come up with the null and alternative hypothesis

What will be the degrees of freedom?

Try to write the test equation


## Paired Samples

Used when we have a predicted mean ($\mu_0$) we would like to assess a single sample distribution from

$$H_0 : \text{Mean}(X_1-X_2) = \bar{d} = 0 \text{    OR    } \mu_1 - \mu_2 = \mu_d = 0$$

$$H_a : \text{Mean}(X_1-X_2) = \bar{d} \neq 0 \text{    OR    } \mu_1 - \mu_2 = \mu_d \neq 0$$

$$df = n_1 - 1$$

$$ t = \frac{\bar{d} - 0}{s_{\bar{d}}} = \frac{\bar{d}}{\frac{s_d}{\sqrt{n}}}$$

## JAMOVI example

Will you get better at the Dinosaur Jumping game with a little practice?

1) Turn your phone to airplane mode

2) Open the chrome web browser and search for ibm.com

3) Play the Dinosaur Jumping game once and record the score you got on paper

4) Continue to play the Jumping game until I call time, when I call time record your final score

5) Turn airplone mode off and use this QR code to enter the initial score (Score_0), and the final score (Score_1)

## JAMOVI example

```{r qr_paired}
code_link <- "https://forms.gle/cMiRFkkjiArtTvkVA"
plot(qr_code(code_link))
print(code_link)
```


